The Calibration Reference Data System

### 2016-02-27 docs/paper/response.txt

Paper section 3 item 1:
The main point being made was that one team may submit a change
to be tested (which could take a while), followed by another team
that has a  more urgent need for an update and tesing, yet must
await the previous test to complete. The teams operate asynchronously
so one test system isn't necessarily sufficient. But the criticism 
is a good one, and given this isn't a very important point in the
paper, we have shortened this issue to say essentially that the
test system didn't address all test needs.

page 6, substitutions.
The wording suggestions are indeed much better and have been 
made.

page 6, last paragraph
Again the suggested text is much better and has been used.

Section 3 comments
------------------

Figures 1-6 have been boxed as requested (we do agree that this
is a signficant improvement)

Figures 7-12 xxxx

Section 4 comments
------------------

These have all been fixed (we hope :-) We thank the reviewer for
taking the time to detail these. After viewing the same or similar
text repeatedly it is quite hard to look as closely as one would
wish, so this is quite helpful.

@@

1) box text file figures
2) improve screen shots

===================================================

Reviewer 2

Section 2
We have reworded to be more explicit that all possible combinations of parameters
must be represented as distinct rows.

Section 3 point 2: "Could an 'is valid' boolean attribute in the database
help out?"
Not really. The file itself may be perfectly good, it is just associated with
the wrong value of parameters. Possbily if coupled with a resubmission with the
right values, but note that may mean many entries have to be so marked, and 
replicated. But as implemented, it didn't have that option.

Section 5: first sentence.
Fixed.
Section 5 submission typo
Fixed
Section 5 point 3
todo****xxxx

Section 6.3 paragraph 2
reworded
Section 6.3 par 4
Made it clear that it is the version of the calibration pipeline that is 
being referred to.
Section 6.4.1 Relational Expressions typo
Fixed
Section 6.4.1 next to last par
Joined two sentences properly.
Section 6.5 par 3
Fixed
Section 7 par 2 punctuation
Fixed
Section 7.2 par 2 begains 
Fixed

### 2016-02-27 docs/paper/crds.tex
boxed figures 1-6 as reviewer 1 requested
git-svn-id: https://aeon.stsci.edu/ssb/svn/crds_server/trunk@2240 37088bc5-bd3f-4c5b-b9de-119bb1e8f7e3
reviewer 2 minor comments
git-svn-id: https://aeon.stsci.edu/ssb/svn/crds_server/trunk@2241 37088bc5-bd3f-4c5b-b9de-119bb1e8f7e3

%% The amssymb package provides various useful mathematical symbols
% \usepackage{amssymb}
\usepackage{gensymb}
%% The amsthm package provides extended theorem environments
%% \usepackage{amsthm}

@@

The most critical aspect of the validation requirements for reference files
concerned the keyword values used to match datasets to the appropriate
references.  Unlike dataset parameter keywords that describe a specific
instrument configuration for one observation, reference file keywords
circumscribe a range of parameter values for which the reference file applies.
In CDBS, ranges of parameter values were often represented by special values
that expand into the discrete values used to define specific instrument
configurations.  As part of submitting reference files, these special values
were combinatorially exploded into multiple database rows that represented each
possible combination of parameter values intended to match to the specified
reference file.  As an example, a dataset APERTURE keyword would describe the
specific aperture an instrument was configured to use for one observation. In
contrast, the same APERTURE keyword specified in a reference file would
describe, indirectly, the combined set of different specific APERTUREs that
reference file was intended to be used for.  When multiple parameters were used
to match datasets to references, the number of distinct rows in the CDBS
database was equal to the product of the number of specific values supported by
a particular reference file for each parameter.

@@

The software for the Calibration Reference Data System (CRDS) 
can be considered to be in one of three categories:

\begin{enumerate}
\item Supporting the bestref functionality.
\item Supporting the submittion of new reference files and all the issues that
go with that.
\item Supporting format, semantic, and parameter constraint checking for both
references and rules files.  Having the capability to detect typos, duplicate 
lines, and grammatical errors in rules files. 
Having the capability to analyze basic properties
such as duplicate or deleted rows,  by pseudo-mode, for supported tables files.
\item Providing utilities for making CRDS useful for operations, instrument
scientists that must generate and monitor reference files, and end users.
\end{enumerate}

@@

The header permits attributes that allow specialized handling of the metadata 
the dataset, including indicating when parameters are relevant or not (e.g., only 
care about its value if the value of another parameter is given as specified), 
how to handle cases where no match is found, and hooks to allow special preprocessing
of parameter values. See the documentation for details on these cases.
\footnote{\url{https://jwst-crds.stsci.edu/static/users\_guide/rmap_syntax.html\#active-header-fields}}
Most of these special features were added to handle the existing complexitites present in the CDBS rule system; we expect these will be used far less for JWST.

@@

\begin{figure*}
\begin{verbatim}
header = {
    'derived_from' : 'jwst_miri_dark_0007.rmap',
    'filekind' : 'DARK',
    'instrument' : 'MIRI',
    'mapping' : 'REFERENCE',
    'name' : 'jwst_miri_dark_0009.rmap',
    'observatory' : 'JWST',
    'parkey' : (('META.INSTRUMENT.DETECTOR', 'META.EXPOSURE.READPATT', 'META.SUBARRAY.NAME'),),
    'sha1sum' : 'bd99f4960096de9ad39a9ce506d5ff985847c4a3',
}

selector = Match({
    ('MIRIFULONG', 'FAST', 'FULL') : UseAfter({
        '2015-01-27 12:05:34' 'jwst_miri_dark_0029.fits',
        '2015-03-22 00:29:01' 'jwst_miri_dark_0032.fits',
        '2015-04-07 09:14:00' 'jwst_miri_dark_0047.fits',
        ...
        }),
    ('MIRIFUSHORT', 'SLOW', 'FULL') : UseAfter({
        '2015-01-25 12:06:34' 'jwst_miri_dark_0099.fits',
        '2015-03-21 00:28:01' 'jwst_miri_dark_0101.fits',
        '2015-04-08 09:13:00' 'jwst_miri_dark_0127.fits',
        ...
        }),
        ...
})
\end{verbatim}
\caption{Example showing basic parameter matching rules coupled with date dependency.}
\label{fig:example1}
\end{figure*}

Figure~\ref{fig:example2} shows the use of the software selection mechanism (for brevity, the header is omitted in this and subsequent examples). In this case if the software version is less than 3.1 (using the typical version semantics) then the first file will be chosen, if less than 5 the second will be chosen, otherwise the last file will be used.

\begin{figure*}
\begin{verbatim}
selector = SelectVersion({
   '<3.1':    'jwst_miri_flat_0065.fits',
   '<5':      'jwst_miri_flat_0073.fits',
   'default': 'jwst_miri_flat_123.fits',
})
\end{verbatim}
\caption{Example showing use of the software version selection mechanism.}
\label{fig:example2}
\end{figure*}

@@

\begin{figure*}
\begin{verbatim}
selector = Bracket({
    1.2: "cref_flatfield_120.fits",
    1.5: "cref_flatfield_124.fits",
    5.0: "cref_flatfield_137.fits",
})
\end{verbatim}
\caption{Example showing bracketing selection.}
\label{fig:example3}
\end{figure*}

@@

\begin{figure*}
\begin{verbatim}
selector = Match({
    ('WFC', 'N/A', 'BETWEEN 1220 1530', 'FR782N') :  "hst_wfc3_flatfile_0027.fits",
    ('WFC', 'N/A', 'BETWEEN 1530 1690', 'FR782N') :  "hst_wfc3_flatfile_0028.fits",
    ...
})
\end{verbatim}
\caption{An example showing matching specified ranges.}
\label{fig:example4}
\end{figure*}

@@

\begin{description}
\item[Lists of Allowable Alternatives:] Allowable values for parameteters
may be supplied as an \textquotesingle OR\textquotesingle
{\tt }list using {\tt |} as a separator. For example:
\texttt {({\ttfamily"}filter1|filter2{\ttfamily"},...)} will allow either value to match the first
parameter.
\item[Wild Cards:] Using an {\tt *} in a string applies familiar wildcard 
matching rules to the supplied string. For example 
\texttt {({\ttfamily"}F*122{\ttfamily"})} will
match any value starting with {\tt F} and ending with {\tt 122}
\item[Regular Expressions:] This allows matching true regular expressions,
allowing for more complex matching rules than simple wildcards permit.
\item[Literal Expressions:] A literal expression is bracketed with curly
braces {\tt \{ \}} and allows inclusion of special characters without 
special interpretation by CRDS, such as {\tt | and \*}.
\item[Relational Expressions:] Relational expressions are braketed by {\tt \#}.
These do numerical comparisons on the supplied parameter value to determine
a match. Supported comparison operators are: 
\texttt {\textgreater, \textgreater =, \textless, \textless =, == and or}. 
For example \texttt {({\ttfamily"}\# >=1 and <37",)} 
will match any number equal to or 
greater than 1 and less than 37.
\item[Between:] This is simply a more convenient way of expressing a range
than using relational expressions. E.g., equivalent to the previous example
is: \texttt{{\ttfamily"}between 1 37{\ttfamily"}}. It is inclusive of the first number.
\item[N/A:] This is equivalent to using a wildcard for the value, but with 
the difference of how weighting is used (discussed below). 
\item[NOT Expressions:] A negation of a match. 
\texttt {{\ttfamily"}not HRC{\ttfamily"}} 
is a match if the value does not match {\tt HRC}.
\item[Substitutions:] Substitutions can be defined in the header. Essentially
they result in the replacement of a string in a rules file with a different
string. This is useful in the case where the string is complex and must be 
entered in many different rules (e.g., it is a combination of many different 
allowable values), or where it is more mnemonic to those that must maintain
the rules than the actual header values are.
\end{description}

With all these match options it is possible to write multiple items that 
match a given case. In order to allow effective defaults for general cases
while allowing specific cases to override the general case, special rules
for disambiguation in these situations have been developed. The basic idea
is that. The precedence is handled by determining the specificity of the
match. Use of specific values, wildcards, or regular expressions increase
the specificity index, while use of N/A does not. If multiple rules match,
the one with the highest specificity index is selected

@@

\begin{figure*}
\begin{verbatim}
header = {
    'derived_from' : 'jwst_nircam_0015.imap',
    'instrument' : 'NIRCAM',
    'mapping' : 'INSTRUMENT',
    'name' : 'jwst_nircam_0016.imap',
    'observatory' : 'JWST',
    'parkey' : ('REFTYPE',),
    'sha1sum' : '0fd29b57f226f8c53bf3dfa7564eccd4e3fc56bf',
}

selector = {
    'DARK' : 'jwst_nircam_dark_0003.rmap',
    'DISTORTION' : 'N/A',
    'DRIZPARS' : 'jwst_nircam_drizpars_0000.rmap',
    'FLAT' : 'jwst_nircam_flat_0000.rmap',
    'GAIN' : 'jwst_nircam_gain_0001.rmap',
    'IPC' : 'jwst_nircam_ipc_0000.rmap',
    'LASTFRAME' : 'N/A',
    'LINEARITY' : 'jwst_nircam_linearity_0003.rmap',
    'MASK' : 'jwst_nircam_mask_0002.rmap',
    'PHOTOM' : 'jwst_nircam_photom_0002.rmap',
    'READNOISE' : 'jwst_nircam_readnoise_0001.rmap',
    'REGIONS' : 'N/A',
    'RESET' : 'N/A',
    'SATURATION' : 'jwst_nircam_saturation_0002.rmap',
    'SPECWCS' : 'N/A',
    'SUPERBIAS' : 'jwst_nircam_superbias_0003.rmap',
    'WCSREGIONS' : 'N/A',
}
\end{verbatim}
\caption{An example of an instrument contex file (.imap)}
\label{fig:example_instrument_context}
\end{figure*}

@@

\begin{figure*}
\begin{verbatim}
header = {
    'derived_from' : 'jwst_0068.pmap',
    'description' : 'Hand edited to support updated .imaps for N/A',
    'mapping' : 'PIPELINE',
    'name' : 'jwst_0069.pmap',
    'observatory' : 'JWST',
    'parkey' : ('META.INSTRUMENT.NAME',),
    'sha1sum' : '691d78053ef710f31bfd8f5bea8ffdc3c1dec1aa',
}

selector = {
    'FGS' : 'jwst_fgs_0016.imap',
    'MIRI' : 'jwst_miri_0043.imap',
    'NIRCAM' : 'jwst_nircam_0016.imap',
    'NIRISS' : 'jwst_niriss_0024.imap',
    'NIRSPEC' : 'jwst_nirspec_0026.imap',
}
\end{verbatim}
\caption{An example of a pipeline context file (.pmap)}
\label{fig:example_pipeline_context}
\end{figure*}

This hierarchy is designed to match that for JWST and as such is currently
designed around the current levels of hierarchy. A possible generalization
of this scheme would allow arbitrary levels of heirarchy, perhaps
different in number depending on the branch taken. Achieving this would
take some modification of the bestref client functionality but is possible
without a large effort. Doing the same for the submission machinery would 
be significantly more work.

>>>>

\usepackage{mdframed}
%% The amssymb package provides various useful mathematical symbols
% \usepackage{amssymb}
\usepackage{gensymb}
%% The amsthm package provides extended theorem environments
%% \usepackage{amsthm}

@@

The most critical aspect of the validation requirements for reference files
concerned the keyword values used to match datasets to the appropriate
references.  Unlike dataset parameter keywords that describe a specific
instrument configuration for one observation, reference file keywords
circumscribe a range of parameter values for which the reference file applies.
In CDBS, ranges of parameter values were often represented by special values
that expand into the discrete values used to define specific instrument
configurations.  As part of submitting reference files, these special values
were combinatorially exploded into multiple database rows that represented each
possible combination of parameter values intended to match to the specified
reference file.  As an example, a dataset APERTURE keyword would describe the
specific aperture an instrument was configured to use for one observation. In
contrast, the same APERTURE keyword specified in a reference file would
describe, indirectly, the combined set of different specific APERTUREs that
reference file was intended to be used for.  When multiple parameters were used
to match datasets to references, the number of distinct rows in the CDBS
database was equal to the number of all possible combinations of parameters.

@@

The software for the Calibration Reference Data System (CRDS) 
can be considered to be in one of four categories:

\begin{enumerate}
\item Supporting the bestref functionality.
\item Supporting the submission of new reference files and all the issues that
go with that.
\item Supporting format, semantic, and parameter constraint checking for both
references and rules files.  Having the capability to detect typos, duplicate 
lines, and grammatical errors in rules files. 
Having the capability to analyze basic properties
such as duplicate or deleted rows,  by pseudo-mode, for supported tables files.
\item Providing utilities for making CRDS useful for operations, instrument
scientists that must generate and monitor reference files, and end users.
\end{enumerate}

@@

The header allows inclusion of attributes that enable specialized handling of the metadata 
of the dataset, including indicating when parameters are relevant or not (e.g., only 
care about its value if the value of another parameter is given as specified), 
how to handle cases where no match is found, and hooks to allow special preprocessing
of parameter values. See the documentation for details on these cases.
\footnote{\url{https://jwst-crds.stsci.edu/static/users\_guide/rmap_syntax.html\#active-header-fields}}
Most of these special features were added to handle the existing complexitites present in the CDBS rule system; we expect these will be used far less for JWST.

@@

\begin{figure*}
\begin{mdframed}
\begin{verbatim}
header = {
    'derived_from' : 'jwst_miri_dark_0007.rmap',
    'filekind' : 'DARK',
    'instrument' : 'MIRI',
    'mapping' : 'REFERENCE',
    'name' : 'jwst_miri_dark_0009.rmap',
    'observatory' : 'JWST',
    'parkey' : (('META.INSTRUMENT.DETECTOR', 'META.EXPOSURE.READPATT', 'META.SUBARRAY.NAME'),),
    'sha1sum' : 'bd99f4960096de9ad39a9ce506d5ff985847c4a3',
}

selector = Match({
    ('MIRIFULONG', 'FAST', 'FULL') : UseAfter({
        '2015-01-27 12:05:34' 'jwst_miri_dark_0029.fits',
        '2015-03-22 00:29:01' 'jwst_miri_dark_0032.fits',
        '2015-04-07 09:14:00' 'jwst_miri_dark_0047.fits',
        ...
        }),
    ('MIRIFUSHORT', 'SLOW', 'FULL') : UseAfter({
        '2015-01-25 12:06:34' 'jwst_miri_dark_0099.fits',
        '2015-03-21 00:28:01' 'jwst_miri_dark_0101.fits',
        '2015-04-08 09:13:00' 'jwst_miri_dark_0127.fits',
        ...
        }),
        ...
})
\end{verbatim}
\end{mdframed}
\caption{Example showing basic parameter matching rules coupled with date dependency.}
\label{fig:example1}
\end{figure*}

Figure~\ref{fig:example2} shows the use of the software selection mechanism where a dependency is indicated for the version of the calibration pipeline (for brevity, the header is omitted in this and subsequent examples). In this case if the version of the calibration pipeline is less than 3.1 (using the typical version semantics) then the first file will be chosen, if less than 5 the second will be chosen, otherwise the last file will be used.

\newpage

\begin{figure*}
\begin{mdframed}
\begin{verbatim}
selector = SelectVersion({
   '<3.1':    'jwst_miri_flat_0065.fits',
   '<5':      'jwst_miri_flat_0073.fits',
   'default': 'jwst_miri_flat_123.fits',
})
\end{verbatim}
\end{mdframed}
\caption{Example showing use of the software version selection mechanism.}
\label{fig:example2}
\end{figure*}

@@

\begin{figure*}
\begin{mdframed}
\begin{verbatim}
selector = Bracket({
    1.2: "cref_flatfield_120.fits",
    1.5: "cref_flatfield_124.fits",
    5.0: "cref_flatfield_137.fits",
})
\end{verbatim}
\end{mdframed}
\caption{Example showing bracketing selection.}
\label{fig:example3}
\end{figure*}

@@

\begin{figure*}
\begin{mdframed}
\begin{verbatim}
selector = Match({
    ('WFC', 'N/A', 'BETWEEN 1220 1530', 'FR782N') :  "hst_wfc3_flatfile_0027.fits",
    ('WFC', 'N/A', 'BETWEEN 1530 1690', 'FR782N') :  "hst_wfc3_flatfile_0028.fits",
    ...
})
\end{verbatim}
\end{mdframed}
\caption{An example showing matching specified ranges.}
\label{fig:example4}
\end{figure*}

@@

\begin{description}
\item[Lists of Allowable Alternatives:] Allowable values for parameteters
may be supplied as an \textquotesingle OR\textquotesingle
{\tt }list using {\tt |} as a separator. For example:
\texttt {({\ttfamily"}filter1|filter2{\ttfamily"},...)} will allow either value to match the first
parameter.
\item[Wild Cards:] Using an {\tt *} in a string applies familiar wildcard 
matching rules to the supplied string. For example 
\texttt {({\ttfamily"}F*122{\ttfamily"})} will
match any value starting with {\tt F} and ending with {\tt 122}
\item[Regular Expressions:] This allows matching true regular expressions,
allowing for more complex matching rules than simple wildcards permit.
\item[Literal Expressions:] A literal expression is bracketed with curly
braces {\tt \{ \}} and allows inclusion of special characters without 
special interpretation by CRDS, such as {\tt | and \*}.
\item[Relational Expressions:] Relational expressions are bracketed by {\tt \#}.
These do numerical comparisons on the supplied parameter value to determine
a match. Supported comparison operators are: 
\texttt {\textgreater, \textgreater =, \textless, \textless =, == and or}. 
For example \texttt {({\ttfamily"}\# >=1 and <37",)} 
will match any number equal to or 
greater than 1 and less than 37.
\item[Between:] This is simply a more convenient way of expressing a range
than using relational expressions. E.g., equivalent to the previous example
is: \texttt{{\ttfamily"}between 1 37{\ttfamily"}}. It is inclusive of the first number.
\item[N/A:] This is equivalent to using a wildcard for the value, but with 
the difference of how weighting is used (discussed below). 
\item[NOT Expressions:] A negation of a match. 
\texttt {{\ttfamily"}not HRC{\ttfamily"}} 
is a match if the value does not match {\tt HRC}.
\item[Substitutions:] Substitutions can be defined in the header. Essentially
they result in the replacement of a string in a rules file with a different
string. This is useful in the case where the string is complex and must be 
entered in many different rules (e.g., it is a combination of many different 
allowable values), or where it is more mnemonic to those that must maintain
the rules than the actual header values are.
\end{description}

With all these match options it is possible to write multiple items that 
match a given case. In order to allow effective defaults for general cases
while allowing specific cases to override the general case, special rules
for disambiguation in these situations have been developed. The basic idea
is that the precedence is handled by determining the specificity of the
match. Use of specific values, wildcards, or regular expressions increase
the specificity index, while use of N/A does not. If multiple rules match,
the one with the highest specificity index is selected

@@

\begin{figure*}
\begin{mdframed}
\begin{verbatim}
header = {
    'derived_from' : 'jwst_nircam_0015.imap',
    'instrument' : 'NIRCAM',
    'mapping' : 'INSTRUMENT',
    'name' : 'jwst_nircam_0016.imap',
    'observatory' : 'JWST',
    'parkey' : ('REFTYPE',),
    'sha1sum' : '0fd29b57f226f8c53bf3dfa7564eccd4e3fc56bf',
}

selector = {
    'DARK' : 'jwst_nircam_dark_0003.rmap',
    'DISTORTION' : 'N/A',
    'DRIZPARS' : 'jwst_nircam_drizpars_0000.rmap',
    'FLAT' : 'jwst_nircam_flat_0000.rmap',
    'GAIN' : 'jwst_nircam_gain_0001.rmap',
    'IPC' : 'jwst_nircam_ipc_0000.rmap',
    'LASTFRAME' : 'N/A',
    'LINEARITY' : 'jwst_nircam_linearity_0003.rmap',
    'MASK' : 'jwst_nircam_mask_0002.rmap',
    'PHOTOM' : 'jwst_nircam_photom_0002.rmap',
    'READNOISE' : 'jwst_nircam_readnoise_0001.rmap',
    'REGIONS' : 'N/A',
    'RESET' : 'N/A',
    'SATURATION' : 'jwst_nircam_saturation_0002.rmap',
    'SPECWCS' : 'N/A',
    'SUPERBIAS' : 'jwst_nircam_superbias_0003.rmap',
    'WCSREGIONS' : 'N/A',
}
\end{verbatim}
\end{mdframed}
\caption{An example of an instrument contex file (.imap)}
\label{fig:example_instrument_context}
\end{figure*}

@@

\begin{figure*}
\begin{mdframed}
\begin{verbatim}
header = {
    'derived_from' : 'jwst_0068.pmap',
    'description' : 'Hand edited to support updated .imaps for N/A',
    'mapping' : 'PIPELINE',
    'name' : 'jwst_0069.pmap',
    'observatory' : 'JWST',
    'parkey' : ('META.INSTRUMENT.NAME',),
    'sha1sum' : '691d78053ef710f31bfd8f5bea8ffdc3c1dec1aa',
}

selector = {
    'FGS' : 'jwst_fgs_0016.imap',
    'MIRI' : 'jwst_miri_0043.imap',
    'NIRCAM' : 'jwst_nircam_0016.imap',
    'NIRISS' : 'jwst_niriss_0024.imap',
    'NIRSPEC' : 'jwst_nirspec_0026.imap',
}
\end{verbatim}
\end{mdframed}
\caption{An example of a pipeline context file (.pmap)}
\label{fig:example_pipeline_context}
\end{figure*}

This hierarchy is designed to match that for JWST and as such is currently
designed around the current levels of hierarchy. A possible generalization
of this scheme would allow arbitrary levels of hierarchy, perhaps
different in number depending on the branch taken. Achieving this would
take some modification of the bestref client functionality but is possible
without a large effort. Doing the same for the submission machinery would 
be significantly more work.
